---
title: "final_project.Rmd"
author: "Scott Mahon"
date: "4/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rvest)
library(xml2)
library(readxl)
library(janitor)
library(ggrepel)
library(ggimage)
library(RCurl)
library(gt)
library(broom)
options(scipen=999)
```
```{r load in wins}

# As of right now, I'm working on the code that ALyssa gave my to find an easier
# way to scrape data from websites. For this week, I manually added the runs per
# team per season as well as strikeouts as a pitching staff to excel files,
# which wasn't too difficult, but we're looking for an easier way to do
# attendance because that would be very difficult to do manually. Also, I'm
# trying to find a way to include team logos instead of dots.

wins1 <- read_excel("raw-data/wins.xlsx") %>%
  filter(!is.na(ABBR))
 wins <- wins1 %>%
  pivot_longer(-ABBR, names_to = "year", values_to = "wins")

```
```{r load in batting average}
battingaverage1 <- read_excel("raw-data/MLBbattingaverage.xlsx") 
 battingaverage <- battingaverage1 %>%
   select (-Name) %>%
  pivot_longer(-ABBR, names_to = "year", values_to = "battingaverage") 
```
```{r load in errors}
errors1 <- read_excel("raw-data/MLBerrors.xlsx")
errors <- errors1 %>%
   select (-Team) %>%
  pivot_longer(-ABBR, names_to = "year", values_to = "errors") 
```
```{r}
runs1 <- read_excel("raw-data/MLBRuns.xlsx")
runs <- runs1 %>%
   select (-Name) %>%
  pivot_longer(-ABBR, names_to = "year", values_to = "runs") 
```
```{r}
Strikeouts_hitters <- read_excel("raw-data/strikeouts_hitters.xlsx") %>%
    select (-Name) %>%
    pivot_longer(-ABBR, names_to = "year", values_to = "Strikeouts_hitters") 
```

```{r}
strikeouts <- read_excel("raw-data/strikeoutsper9.xlsx") %>%
   select (-Name) %>%
  pivot_longer(-ABBR, names_to = "year", values_to = "strikeouts") 
```
```{r}
StolenBases <- read_excel("raw-data/StolenBases.xlsx") %>%
    pivot_longer(-ABBR, names_to = "year", values_to = "StolenBases") 
```
```{r}
Walks <- read_excel("raw-data/Walks.xlsx") %>%
    pivot_longer(-ABBR, names_to = "year", values_to = "Walks") 
```

```{r}
attendance <- read_excel("raw-data/MLBattendance.xlsx") %>%
  select(-Name) %>%
  select(-max) %>%
  pivot_longer(-ABBR, names_to = "year", values_to = "attendance") 
```
```{r}
mlblogos <- read_excel("raw-data/mlblogosfinal.xlsx")
```
```{r}
AllStars <- read_excel("raw-data/allstars.xlsx") %>%
    pivot_longer(-ABBR, names_to = "year", values_to = "AllStars") 
```

```{r join tables}
joined <- wins %>%
  full_join (battingaverage, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(errors, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(runs, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(strikeouts, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(attendance, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(Strikeouts_hitters, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(Walks, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(StolenBases, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(AllStars, by = c("ABBR" = "ABBR", "year" = "year")) %>%
  full_join(mlblogos, by = c("ABBR" = "ABBR"))
```

```{r}
joined %>%
  filter (year == "2019") %>%
  ggplot(aes(x = attendance, y = wins)) +
  geom_image(aes(image = URL), size = 0.05)
```
```{r}
joined %>%
        lm(scale(wins) ~ scale(battingaverage) + scale(errors) + scale(attendance) + scale(runs) + scale(Strikeouts_hitters) + scale(strikeouts),
           data = .) %>%
        tidy(conf.int = TRUE) %>%
        select(term, estimate, conf.low, conf.high) %>%
        rename(
            "Variable" = term,
            "Estimate" = estimate,
            "Lower Bound" = conf.low,
            "Upper Bound" = conf.high) %>%
            mutate_if(is.numeric, round, 4) %>%
        gt()
```
```{r}
joined %>%
        lm(scale(wins) ~ scale(AllStars),
           data = .) %>%
        tidy(conf.int = TRUE) %>%
        select(term, estimate, conf.low, conf.high) %>%
        rename(
            "Variable" = term,
            "Estimate" = estimate,
            "Lower Bound" = conf.low,
            "Upper Bound" = conf.high) %>%
            mutate_if(is.numeric, round, 4) %>%
        gt()
```
```{r}
joined %>%
            lm(scale(wins) ~ scale(battingaverage)*scale(runs), 
               data = .) %>%
            tidy(conf.int = TRUE) %>%
            select(term, estimate, conf.low, conf.high) %>%
            rename(
                "Variable" = term,
                "Estimate" = estimate,
                "Lower Bound" = conf.low,
                "Upper Bound" = conf.high) %>%
            mutate_if(is.numeric, round, 4) %>%
            gt()
```


```{r}
# joined1 <- joined %>%
# filter(year == 2018) %>%
# cor(joined1$battingaverage, joined1$wins) 
joined1 <- joined %>% 
  select(wins, battingaverage, errors, runs, strikeouts, attendance) 
joined2 <- scale(joined1, center = TRUE, scale = TRUE) 

joined %>% 
  group_by(year) %>%
        lm(wins ~ attendance, data = .) %>%
        tidy(conf.int = TRUE) %>% 
        select(term, estimate, conf.low, conf.high) %>%
        rename(
            "Variable" = term,
            "Estimate" = estimate,
            "Lower Bound" = conf.low,
            "Upper Bound" = conf.high
        ) %>%
        gt()
```
```{r}
          joined %>%
            filter(ABBR == "CLE") %>%
            ggplot(aes(x = year, y = battingaverage)) +
            geom_point() + theme_classic()
```

```{r}

```

I have five data sets loaded in and have added more graphs, so I now have 31
different graphs, one for runs, batting average, strikeouts per 9 innings as
pitchers and errors for the years 2010-2019 (strikeouts I only have 2019). My
next steps are to continue adding data, most importantly attendance, as well as
some others. I also want to begin exploring data that mightn ot exist explicitly
such as number of offseason transactions, which I would most likely have to just
manually count. I also want to add images in place of the points on the plots
and also see if there is a way to denote which team won the world series each
year. I have already added lines of best fit, and hope to use the linear
regressions to show something, too although I'm not quite sure what that is
going to be at this point. 

```{r}
attendance <-joined %>%
  filter (year == "2019") %>%
  ggplot(aes(x = attendance, y = wins)) +
  geom_image(aes(image = URL), size = 0.05) +
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "attendance", y = "wins", 
      title = "Wins vs attendance in 2019") 

attendance + geom_smooth(method =lm, se = FALSE, color = "black") 
```

```{r graph errors 2019}
errors2019 <-joined %>%
  filter (year == "2019") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2019") 

errors2019 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2018}
errors2018 <- joined %>%
  filter (year == "2018") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2018") 

errors2018 + geom_smooth(method = lm, se = FALSE, color = "black") 
```
```{r graph errors 2017}
errors2017 <- joined %>%
  filter (year == "2017") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2017") 

errors2017 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2016}
errors2016 <- joined %>%
  filter (year == "2016") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2016") 

errors2016 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2015}
errors2015 <- joined %>%
  filter (year == "2015") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2015") 

errors2015 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2014}
errors2014 <- joined %>%
  filter (year == "2014") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2014") 

errors2014 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2013}
errors2013 <- joined %>%
  filter (year == "2013") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2013") 

errors2013 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2012}
errors2012 <- joined %>%
  filter (year == "2012") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2012") 

errors2012 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2011}
errors2011 <- joined %>%
  filter (year == "2011") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2011") 

errors2011 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph errors 2010}
errors2010 <- joined %>%
  filter (year == "2010") %>%
  ggplot(aes(x = errors, y = wins, color = ABBR)) +
  geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "errors", y = "wins", 
      title = "Wins vs Errors in 2010") 

errors2010 + geom_smooth(method =lm, se = FALSE, color = "black") 
```


```{r graph batting average 2019}
BA2019 <- joined %>%
  filter (year == "2019") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2019") 

BA2019 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2018}
BA2018 <- joined %>%
  filter (year == "2018") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2018") 

BA2018 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2017}
BA2017 <- joined %>%
  filter (year == "2017") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2017") 

BA2017 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2016}
BA2016 <- joined %>%
  filter (year == "2016") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2016") 

BA2016 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2015}
BA2015 <- joined %>%
  filter (year == "2015") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2015") 

BA2015 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2014}
BA2014 <- joined %>%
  filter (year == "2014") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2014") 

BA2014 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2013}
BA2013 <- joined %>%
  filter (year == "2013") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2013") 

BA2013 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2012}
BA2012 <- joined %>%
  filter (year == "2012") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2012") 

BA2012 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2011}
BA2011 <- joined %>%
  filter (year == "2011") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2011") 

BA2011 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph batting average 2010}
BA2010 <- joined %>%
  filter (year == "2010") %>%
  ggplot(aes(x = battingaverage, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "batting average", y = "wins", 
      title = "Wins vs Batting Average in 2010") 

BA2010 + geom_smooth(method =lm, se = FALSE, color = "black") 
```


```{r graph runs 2019}
runs2019 <- joined %>%
  filter (year == "2019") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2019") 

runs2019 +   
  geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2018}
runs2018 <- joined %>%
  filter (year == "2018") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2018") 

runs2018 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2017}
runs2017 <- joined %>%
  filter (year == "2017") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2017") 

runs2017 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2016}
runs2016 <- joined %>%
  filter (year == "2016") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2016") 

runs2016 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2015}
runs2015 <- joined %>%
  filter (year == "2015") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2015") 

runs2015 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2014}
runs2014 <- joined %>%
  filter (year == "2014") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2014") 

runs2014 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2013}
runs2013 <- joined %>%
  filter (year == "2013") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2013") 

runs2013 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2012}
runs2012 <- joined %>%
  filter (year == "2012") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2012") 

runs2012 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2011}
runs2011 <- joined %>%
  filter (year == "2011") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2011") 

runs2011 + geom_smooth(method =lm, se = FALSE, color = "black") 
```
```{r graph runs 2010}
runs2010 <- joined %>%
  filter (year == "2010") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Runs in 2010") 

runs2010 + geom_smooth(method =lm, se = FALSE, color = "black") 
```

```{r graph strikeoutsper9 2019}
strikeoutsper9_2019 <- joined %>%
  filter (year == "2019") %>%
  ggplot(aes(x = runs, y = wins, color = ABBR)) +
    geom_point() + 
  geom_text_repel(aes(label=ABBR), size = 3) +
  labs(x = "runs", y = "wins", 
      title = "Wins vs Strikeouts Against hitters in 2019") 

strikeoutsper9_2019 + geom_smooth(method =lm, se = FALSE, color = "black") 
```

```{r, include=FALSE}
  rankings <- read_html("https://www.teamrankings.com/mlb/stat/errors-per-game")
  html_text(rankings)
  tab <- rankings %>% html_nodes("table")
  tab <- tab[[1]] %>% html_table %>%
    select(2, 3, 8)
```